---
title: "FEC_Data"
author: "B. Madhusudan"
date: "May 20, 2016"
output: html_document
---

This is Federal Election Commission data for the Presidential race for 2016;
specifically data for the state of NY; this data was last updated on
21-April-2016

The dataset can be found here: http://fec.gov/disclosurep/PDownload.do
[It's the NY.zip file]

```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

library(GGally)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(Hmisc)
library(xts)
library(scales)

```

Before getting started, this is an important link!
ftp://ftp.fec.gov/FEC/Presidential_Map/2016/DATA_DICTIONARIES/CONTRIBUTOR_FORMAT.txt

This shows what the data elements mean!

Let's load the data & look at all the players

NOTE: In order for read.csv to parse this properly, I needed to append an extra
comma at the end of the row header. If the extra comma wasn't appended, a 
'duplicate row.names' error would have resulted.


```{r Initialize, echo=FALSE}
# Below is an alternate to appending a comma to the row header. The below code
#would use row.names = NULL and then shift the column names over by 1
#fec_data <- read.csv("P00000001-NY.csv", header=TRUE, strip.white=TRUE,
#                     row.names=NULL)
#colnames(fec_data) <- c(colnames(fec_data)[-1],"x")
#fec_data$x <- NULL

fec_data <- read.csv("P00000001-NY.csv", header=TRUE, strip.white=TRUE)

# Let's look at all the players...

unique(fec_data$cand_nm)

```
Interesting! Who is Jill Stein?? Also Gary Johnson and James Gilmore...

Going to get rid of some of the fringe players and less popular candidates

```{r Cleaning Up}
fec_data <- fec_data[fec_data$cand_nm %nin% c("Stein, Jill",
"Webb, James Henry Jr.", "Santorum, Richard J.", "Lessig, Lawrence",
"Gilmore, James S IIII", "Johnson, Gary", "Pataki, George E.", "Paul, Rand",
"Huckabee, Mike","Jindal, Bobby","O\'Malley, Martin Joseph","Fiorina, Carly",
"Perry, James R. (Rick)"),]


fec_data <- as.data.frame(lapply(fec_data, function (x) if (is.factor(x))
                          factor(x) else x)) 
```

Convert the zip into a factor and remove the extra +4 digits
Also save the first 3 digits of the zip separately as that is useful geographic
information [denotes an SCF: Sectional Center Facility]

```{r echo=FALSE}

fec_data$zip <- substr(as.character(fec_data$contbr_zip),1,5)
fec_data$zip <- factor(fec_data$zip)
fec_data$zip_three_dig <- substr(fec_data$zip, 1,3)
fec_data$zip_three_dig <- factor(fec_data$zip_three_dig)

```

Obtaining extra info on the zip codes from 2010 Census Data & then group the
data by SCF and obtain the total population for it
```{r echo=FALSE}

zip_code_data <- read.csv("2010CensusPopulationData.csv", header=TRUE,
                          strip.white=TRUE, row.names=NULL,
                          colClasses = c("character",NA))

three_digit_zip_data <- zip_code_data %>% 
  group_by(scf=substr(zip_code_data$Zip,1,3)) %>% 
  summarise(pop = sum(Population))

# Below is OLD ; useful before I figured out colClasses option in read.csv
#before three_digit_zip_data <- zip_code_data %>% #group_by(scf=substr(as.character(zip_code_data$Zip),1,3))
#%>% summarise(pop = #sum(Population))

three_digit_zip_data$scf <- factor(three_digit_zip_data$scf)
```

Convert dates & get rid of some unnecessary fields
```{r echo=FALSE}

fec_data$contb_receipt_dt <- as.Date(fec_data$contb_receipt_dt,
                                     format="%d-%b-%y")

fec_data$contbr_zip <- NULL # dont need this anymore
fec_data$st <- NULL # don't need this anymore as NY is only state

```

Supplementing the data set with additional attributes of the candidates
including their party, gender, and dates they dropped out of the campaign
```{r echo=FALSE}

# add affiliation [republican or democrat] and gender data
cand_names <- unique(fec_data$cand_nm)

affiliation <- c("D","R","R","R","R", "R", "D","R","R","R", "R")

gender <- c("M","M","M", "M", "M", "M", "F", "M", "M", "M", "M")

drop_out_date <- as.Date(c(NA,"2016-05-03", "2015-09-21", "2016-02-20",
                           "2016-03-15", "2016-02-10", NA, "2015-12-21", NA,
                           "2016-03-04", "2016-05-04"))

df = data.frame(cand_names, affiliation, gender, drop_out_date)
fec_data <- merge(fec_data, df, by.x="cand_nm", by.y="cand_names")

```

Need to use a geom_bar here; geom_histogram does not work because this is not
continuous data.

```{r}

f <- function(DS, x, y, geom, opts=NULL) {
  aes <- eval(substitute(aes(x, y),
    list(x = substitute(x), y = substitute(y))))
  p <- ggplot(DS, aes) + geom + opts
  
}

```

```{r echo=FALSE}

#p <- f(fec_data, cand_nm,(..count..), geom_bar(aes(fill=cand_nm)))

ggplot(data=fec_data,aes(cand_nm)) + geom_bar(aes(fill=cand_nm)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# ggplot(data=fec_data,aes(cand_nm)) + stat_count()
```
<br>
Bernie Sanders & Hillary Clinton have the most # of contributions by far.

```{r echo=FALSE}
summary(fec_data$contb_receipt_dt)
```

Interesting! the first contribution date is back in 2013! to whom, by whom?
Why so early?

```{r Examining the early contributions: pre-2015}

ggplot(data=subset(fec_data, contb_receipt_dt < as.Date("2014-01-01")),
       aes(cand_nm)) + geom_bar(aes(fill=cand_nm)) + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
<br>
Answer: Marco Rubio!

How about the other candidates before 2015? who believed themselves destined
for greatness so early on?

```{r echo=FALSE}

ggplot(data=subset(fec_data, contb_receipt_dt < as.Date("2015-01-01")),
       aes(cand_nm)) + geom_bar(aes(fill=cand_nm)) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r echo=FALSE}
hist(fec_data$contb_receipt_dt, breaks="months")
```

Makes sense that the # of contributions is steadily climbing and hits a
maximum on the latest date we have

Let's look at how much cash came in each year

```{r echo=FALSE}
amounts_by_year <- fec_data %>% group_by(dt=format(contb_receipt_dt,'%Y')) %>%
  summarise(amt = sum(contb_receipt_amt),n=n())

ggplot(data=amounts_by_year, aes(dt, amt)) +
  geom_bar(stat="identity")

# another way of obtaining the data; ;using old-school R; not dplyr
x <- aggregate(fec_data$contb_receipt_amt, 
               by=list(yr = format(fec_data$contb_receipt_dt, '%Y')), "sum")

ggplot(data=x, aes(yr, x)) + geom_bar(stat ="identity")
  
```

2013 & 2014 barely register.
2015 stands tall but 2016 is almost caught up (and this data is only 3 months
into the year!)

What are the actual amounts?

```{r echo=FALSE}
amounts_by_year

```
You can also see that there are already way many more contributions in the
first 3 months of 2016 than there were combined in 2013, 2014, 2015. The Power
of an Election Year!

Pre-2015 amount is miniscule compared to the action in 2015. Further
justification to drop the pre-2015 data.

Going to eliminate data from the data set because I believe they are outliers
or may have some arcane political work-arounds involved

1. Eliminate anything prior to April 1, 2015 because don't believe candidates
for the most part started fully

2. Eliminated any election_tp codes other than "P2016" [so only focused on 
Primary 2016 contributions]

```{r echo=FALSE}

fec_subset <- subset(fec_data, contb_receipt_dt >= as.Date("2015-04-01"))

```

Let's look at the $ by election type; 
```{r echo=FALSE}
aggregate(fec_subset$contb_receipt_amt, by=list(tp=fec_subset$election_tp),
          "sum")

summary(fec_subset$election_tp)
```
-53,770 when it's blank election type? composed of 27 observations

G2016 has $1M ;composed of 1,398 observations

and people are even giving to P2020 election cycle!
(Upon further review, there were only 3 contributions and they were noted as "REDESIGNATIONS"; Interestingly they were all to Lindsey Graham and by prominent
NYers, 2 of whom are married to each other! So my guess is Lindsey had some
sort of party in NY perhaps?)

let's nix the above and only concentrate on P2016 data

```{r echo=FALSE}
fec_P2016 <- subset(fec_subset, election_tp=="P2016")

```

So now we're looking only at contributions made after April 2015 and for
P2016 cycle

Let's again plot the contributions

```{r echo=FALSE}


ggplot(data=fec_P2016,aes(x=cand_nm)) + geom_bar()+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Still shows that Hillary Clinton and Bernie Sanders have gotten the most # of
contributions.

```{r echo=FALSE}

x <- aggregate(fec_P2016$contb_receipt_amt, 
               by=list(party = fec_P2016$affiliation), "sum")

ggplot(data=x, aes(party, x)) + geom_bar(stat ="identity")

```

The Dems have definitely outraised the Republicans.

Let's look now at the amounts that were raised

```{r facet, echo=FALSE, fig.height=7, fig.width=10}

res <- aggregate(fec_P2016$contb_receipt_amt, 
                 by=list(tp=fec_P2016$cand_nm), "sum")

# ALTERNATIVE: aggregate(contb_receipt_amt ~ cand_nm, data=fec_P2016, sum)

res[order(-res$x),]

contribs_by_cand <- fec_P2016 %>% group_by(cand_nm) %>%     summarise(mean_cont=mean(contb_receipt_amt),
median_contrib=median(contb_receipt_amt),
min_contrib=min(contb_receipt_amt),
max_contrib = max(contb_receipt_amt),
sum_contrib = sum(contb_receipt_amt),
n=n())

p1 <- ggplot(data=contribs_by_cand, aes(x=cand_nm, y=mean_cont)) +
geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(data=contribs_by_cand, aes(x=cand_nm, y=median_contrib)) +   geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p3 <- ggplot(data=fec_P2016,aes(x=cand_nm)) + geom_bar() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(p1, p2, p3, ncol=1)

```

This is an interesting perspective here. If both the mean and median are high for
a candidate (e.g. see Chris Christie & Jeb Bush), along with a small # of donors,
this is an indication that they had a concentrated number of people who backed
their campaigns with out-sized contributions.

Meanwhile, Bernie is interesting in that he got both small median & mean
contributions, but because he had so many contributions, he has the 2nd biggest
haul (after Hillary). He really is being powered by the many. Hillary has the edge,
however, because her mean contributions are larger and she also has quite a number of
people contributing.

Look at zip codes: which one contributed most and to whom? As a reminder, we
are looking at the first 3 digits which constitute an SCF

```{r, fig.width=10, fig.height=7,echo=FALSE}

contribs_by_zip <- fec_P2016 %>% group_by(affiliation, zip_three_dig) %>%
  summarise(sum_contrib = sum(contb_receipt_amt),n=n()) 

ggplot(data=contribs_by_zip, aes(x=zip_three_dig, y=sum_contrib)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Sum of Contributions ($)")
```

Very Interesting! The contributions from 2 areas look vastly higher than
any others.

Looking at the data, "100" & "101" greatly surpass any other areas!

```{r echo=FALSE}

highest_zips <- subset(fec_P2016, zip_three_dig %in% c(100, 101)) %>%
  group_by(zip_three_dig, affiliation) %>% summarise(sum_contrib = sum(contb_receipt_amt),n=n())

ggplot(data=highest_zips, aes(x=zip_three_dig, y=sum_contrib, 
                              fill=affiliation)) +     geom_bar(stat="identity",position=position_dodge())

highest_zips <- subset(fec_P2016, zip_three_dig %in% c(100, 101)) %>%
  group_by(zip_three_dig, affiliation, cand_nm) %>% summarise(sum_contrib=sum(contb_receipt_amt),n=n())

ggplot(data=highest_zips, aes(x=zip_three_dig, y=sum_contrib, fill=cand_nm)) + geom_bar(stat="identity",position=position_dodge()) + 
  ylab("Sum of Contributions ($)")

```
<br>
The above shows Hillary has a commanding lead even in the wealthiest zip codes

Let's look at the average contribution per SCF

``` {r, fig.width=10, fig.height=7}
contribs_by_zip <- fec_P2016 %>% group_by(zip_three_dig) %>% summarise(
sum_contrib = sum(contb_receipt_amt),
n=n())

joined_zip_data <- contribs_by_zip %>% left_join(three_digit_zip_data, 
                                                 c("zip_three_dig" = "scf"))

joined_zip_data <- joined_zip_data %>% 
  mutate(avg_cont_per_capita=sum_contrib/pop, 
         avg_cont_per_contributor=sum_contrib/n)

ggplot(data=joined_zip_data, aes(x=zip_three_dig, y=avg_cont_per_capita)) + geom_bar(stat="identity",position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Avg Contribution Per Capita ($)")

```

Definitely 101 takes the cake for average contribution per capita!
Although 100 & 101 have near similar total contributions, 101 has far
fewer people living there.

Actually, based on one reviewer's feedback, there was a suggestion to plot it on a log scale to really be able to compare. This is what is done below and is quite fascinating

```{r, fig.width=10, fig.height=7}

ggplot(data=joined_zip_data, aes(x=zip_three_dig, y=avg_cont_per_capita)) + geom_bar(stat="identity",position=position_dodge()) +
  scale_y_log10(labels = comma) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Avg Contribution Per Capita ($)")

```
<br>
You can see a large # of bars are below 1.00 and some below .01 even! This basically means that there were not a lot of contributions given in those areas. Especially in areas with large numbers of people, those people are not contributing on the whole to the campaigns!

Now, let's compare 'avg contribution per capita' vs. 'avg contribution
per contributor'

```{r, fig.width=10, fig.height=7, echo=FALSE}

joined_zip_data <- joined_zip_data %>% gather(avg_contrib,amt,-1:-4)

ggplot(data=joined_zip_data, aes(x=zip_three_dig, y=amt, fill=avg_contrib)) + geom_bar(stat="identity",position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Amount ($)")

```

Wow, even though 100 & 101 had the most contributions, 

Far & away, Hillary Clinton is getting the most bucks.

```{r echo=FALSE}

contribs_by_job <- fec_P2016 %>% group_by(contbr_occupation) %>%
  summarise(sum_contrib = sum(contb_receipt_amt),n=n())

```

Just for fun, I'm going to look at anyone who's an ACTOR
```{r echo=FALSE}
#Curious about actors!
  
actors <- subset(fec_P2016, contbr_occupation=="ACTOR") %>% 
  group_by(affiliation) %>% 
  summarise(sum_contrib = sum(contb_receipt_amt),n=n())

ggplot(data=actors, aes(x=affiliation, y=sum_contrib)) + geom_bar(stat="identity",position=position_dodge())

```

All Actors have gone to the Democratic side. Not a single actor has
contributed to a Republican candidate!

Now, we will look at contributions on a daily basis for each candidate

```{r, echo=FALSE}

amounts_by_day <- fec_P2016 %>% group_by(contb_receipt_dt, cand_nm) %>%
  summarise(amt = sum(contb_receipt_amt),n=n())
  
  ggplot(data=amounts_by_day, aes(contb_receipt_dt, amt, color=cand_nm)) +
    geom_line() + ylab("Amount ($)")
 
```
<br>
Hillary has some huge spikes! Looking at the top couple of data points though
indicates the reason:  31-March, 29-Feb, 31-Dec. If anyone subscribes to
candidates' mailing lists, this is obvious; There are always HUGE drives to
solicit $$ at the end of the month. But Hillary's machine is way stronger than
everyone else's.

Now, let's look at each of the candidates' hauls over time; this shows the
cumulative sum over time


```{r NEXT, fig.width=10, fig.height=7, echo=FALSE}

cum_amounts_by_day <- subset(fec_P2016[order(fec_P2016$contb_receipt_dt),]) %>% group_by(cand_nm) %>% mutate(cumsum=cumsum(contb_receipt_amt))

cum_amounts_by_day <-
cum_amounts_by_day[order(cum_amounts_by_day$contb_receipt_dt),]

ggplot(data=cum_amounts_by_day, aes(contb_receipt_dt, cumsum, color=cand_nm)) +
  geom_line() +
 scale_x_date(limits=as.Date(c(NA,'2016-04-01')),date_breaks="1 month") +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-01"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-09"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-20"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-23"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-27"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-01"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-08"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-15"))), linetype=2) +
  geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-22"))), linetype=2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Cumulative Amount ($)") + scale_y_continuous(labels=comma)

```

Hillary's climb soars over everyone else

Now let's look at the same data but not on a cumulative basis and only
at Republicans

```{r, echo=FALSE}

amounts_by_day_R <- subset(fec_P2016, affiliation=='R') %>% 
  group_by(contb_receipt_dt, cand_nm) %>% 
  summarise(amt = sum(contb_receipt_amt),n=n())
  
ggplot(data=amounts_by_day_R, aes(contb_receipt_dt, amt, color=cand_nm)) +
geom_line() +
scale_y_continuous(limits=c(0,45000), labels=comma)+
scale_x_date(limits=as.Date(c('2016-01-01','2016-04-01')),
              date_breaks="1 month") +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-01"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-09"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-20"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-23"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-02-27"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-01"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-08"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-15"))), linetype=2) +
geom_vline(aes(xintercept = as.numeric(as.Date("2016-03-22"))), linetype=2) +
ylab("Amount ($)")
```
<br>
This graph is messy. Would really need to select a subset of candidates to
filter on. But I will move on from here.

Let's look at when the first and last contributions were to each candidate with a
vertical line demarcating Jan-01-2016 & additionally the drop out dates 
of each candidate

```{r, fig.width=10, fig.height=7, echo=FALSE}
min_max_contrib_dates <- fec_data %>% group_by(cand_nm, drop_out_date) %>% 
  summarise(first_cont = min(contb_receipt_dt),
            last_cont = max(contb_receipt_dt), n=n())

new_d <- min_max_contrib_dates %>% 
gather(dt_type, dat, -cand_nm, -n, -drop_out_date)

ggplot(data=new_d, aes(dat, cand_nm)) + geom_line() + scale_x_date(limits=as.Date(c(NA,'2016-04-01')),date_breaks="1 month") + 
theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
geom_vline(aes(xintercept = as.numeric(as.Date("2016-01-01"))), linetype=2) +
geom_point(aes(drop_out_date, cand_nm, color='red'))
```

Interesting. Candidates were getting money even AFTER they dropped out!
Scott Walker & Lindsey Graham dropped out in 2015 and they're still
getting contributions!

Let's take a look at amounts raised before & after each candidates 
drop out dates

```{r, fig.width=10, fig.height=7, echo=FALSE}

#before_and_after_dropout <- fec_P2016 %>% 
#group_by(cand_nm, drop_out_date) %>% 
#summarise(sum_total = sum(contb_receipt_dt), sum_contrib_before = #sum(contb_receipt_amt[contb_receipt_dt <= drop_out_date]),n=n())

# before_and_after_dropout <- fec_P2016 %>% 
#group_by(cand_nm, drop_out_date) %>% 
#summarise(sum_total = sum(contb_receipt_dt), sum_contrib_before = #sum(fec_P2016[contb_receipt_dt <= #drop_out_date,'contb_receipt_amt']),n=n())
  
before_and_after_dropout <- fec_P2016 %>% 
group_by(cand_nm, drop_out_date) %>%  
mutate( sum_contrib_before = 
sum(fec_P2016[contb_receipt_dt <= drop_out_date,'contb_receipt_amt'])) %>%
mutate( sum_contrib_after = 
sum(fec_P2016[contb_receipt_dt >   drop_out_date,'contb_receipt_amt']))
  
grouped_data <- before_and_after_dropout %>% distinct(cand_nm) %>%
select(sum_contrib_before, sum_contrib_after)
    
hh <- grouped_data %>% 
gather(period, amount, sum_contrib_before, sum_contrib_after)
  
ggplot(data=hh, aes(x=cand_nm, y=amount, fill=period)) +
geom_bar(stat="identity", position=position_dodge()) + 
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
ylab("Amount ($)")
```


##### EXTRAS #####

A few additional things I was motivated to do after the first Udacity review

## EXTRA 1 ##

This was in the NYTimes on 29-May-2016. 

http://www.nytimes.com/2016/05/29/business/they-tilt-right-but-top-chief-executives-dont-give-to-trump.html?smid=nytcore-iphone-share&smprod=nytcore-iphone

'An analysis of political donation from chief executives shows broad support for
Republican candidates. Except for the presumptive nominee.'

Seems like an ideal thing for me to cross-verify!

```{r Verify_NYT_Article}

top_ppl <- fec_P2016[grep("^(PRESIDENT|CEO|CHIEF)",ignore.case=TRUE, fec_P2016$contbr_occupation),]

where_top_ppl_gave <- top_ppl %>% group_by(cand_nm) %>% 
  summarise(amt_given = sum(contb_receipt_amt),
            n=n())
            
ggplot(data=where_top_ppl_gave, aes(x=cand_nm, y=amt_given)) +
geom_bar(stat="identity", position=position_dodge()) + 
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
ylab("Amount ($)")

titans <- fec_P2016[grep("^(wynn|benioff|hess|murdoch|whitman|wexner|hugin)",  fec_P2016$contbr_nm, ignore.case=TRUE),]
```

Interesting. This seems to contradict the NYT Article. Hillary Clinton is by far
once again the biggest recipient even by all the top ppl
[CEOs, C-Officers, Presidents]!

A few things at play here:

1. NY is a very blue state; I think this is probably the #1 explanation for why
Hillary has gotten more $. The article uses data gathered from the Center for
Responsive Politics which I'm sure used all the states' data combined.

2. A lot of the donors in the article contributed via PAC and organized groups.
The dataset I am using only consists of individual contributors so this is not
an apple to apples comparison.

3. My capturing of the Chief* titles via the grep command may have grabbed other
people that aren't actually top execs so that could have skewed the results;
Also lots of people like using inflated titles or might own small companies, so
this analysis probably caught a lot of people who aren't titans of industry.

4. I also grepped for the names mentioned in the article. The only person that
popped up of signifance was Wendi Murdoch, the ex-wife of Rupert Murdoch and she
gave money to Hillary!

## EXTRA 2 ##

Using Maps! [As suggested by a Udacity Reviewer, I thought I'd give this a go]
Used http://www.computerworld.com/article/3038270/data-analytics/create-maps-in-r-in-10-fairly-easy-steps.html

I'm going to focus in on data in NYC for visualization purposes.

```{r NYC Zips}

# From: https://www.health.ny.gov/statistics/cancer/registry/appendix/neighborhoods.htm

bronx_zips <- c(10453, 10457, 10460,10458, 10467, 10468,10451, 10452, 10456,
10454, 10455, 10459, 10474,10463, 10471,10466, 10469, 10470, 10475,
10461, 10462,10464, 10465, 10472, 10473)
bronx_zips <- as.factor(bronx_zips)

bronx_neighborhoods <- c(rep("Central Bronx", 3),
rep("Bronx Park and Fordham", 3), rep("High Bridge and Morrisania",3),
rep("Hunts Point and Mott Haven",4), rep("Kingsbridge and Riverdale",2),
rep("Northeast Bronx", 4), rep("Southeast Bronx",6))

bronx_data <- data.frame("ZipCode" = bronx_zips,
"Neighborhood"=bronx_neighborhoods)


brooklyn_zips <- c(11212, 11213, 11216, 11233, 11238,11209, 11214, 11228,
11204, 11218, 11219, 11230,	11234, 11236, 11239,	11223, 11224, 11229, 11235,
11201, 11205, 11215, 11217, 11231,11203, 11210, 11225, 11226,11207, 11208,
11211, 11222,11220, 11232,11206, 11221, 11237)
brooklyn_zips <- as.factor(brooklyn_zips)

brooklyn_neighborhoods <- c(rep("Central Brooklyn",5), 
rep("Southwest Brooklyn",3), rep("Borough Park",4), 
rep("Canarsie and Flatlands",3), rep("Southern Brooklyn",4),
rep("Northwest Brooklyn",5), rep("Flatbush",4), 
rep("East New York and New Lots",2),rep("Greenpoint",2), rep("Sunset Park",2),
rep("Bushwick and Williamsburg",3))

bklyn_data <- data.frame("ZipCode" = brooklyn_zips,
"Neighborhood"=brooklyn_neighborhoods)


manhattan_zips <- c(10026, 10027, 10030, 10037, 10039,10001, 10011, 10018, 
10019, 10020, 10036,10029, 10035,10010, 10016, 10017, 10022,10012, 10013, 10014,
10004, 10005, 10006, 10007, 10038, 10280,	10002, 10003, 10009,10021, 10028, 
10044, 10065, 10075, 10128,10023, 10024, 10025,10031, 10032, 10033, 10034, 
10040)
manhattan_zips <- as.factor(manhattan_zips)

manhattan_neighborhoods <- c(rep("Central Harlem",5), 
rep("Chelsea and Clinton",6), rep("East Harlem",2), 
rep("Gramercy Park and Murray Hill",4), rep("Greenwich Village and Soho",3),
rep("Lower Manhattan",6),rep("Lower East Side",3), rep("Upper East Side",6),
rep("Upper West Side",3), rep("Inwood and Washington Heights",5))

manh_data <- data.frame("ZipCode" = manhattan_zips,
"Neighborhood"=manhattan_neighborhoods)

queens_zips <- c(11361, 11362, 11363, 11364,11354, 11355, 11356, 11357, 11358,
11359, 11360, 11365, 11366, 11367,11412, 11423, 11432, 11433, 11434, 11435, 
11436, 	11101, 11102, 11103, 11104, 11105, 11106,11374, 11375, 11379, 11385,
11691, 11692, 11693, 11694, 11695, 11697,11004, 11005, 11411, 11413, 11422, 
11426, 11427, 11428, 11429,	11414, 11415, 11416, 11417, 11418, 11419, 11420, 
11421, 11368, 11369, 11370, 11372, 11373, 11377, 11378)
queens_zips <- as.factor(queens_zips)


queens_neighborhoods <- c(rep("Northeast Queens",4), rep("North Queens",7),
rep("Central Queens",3), rep("Jamaica",7), rep("Northwest Queens",6),
rep("West Central Queens",4), rep("Rockaways",6), rep("Southeast Queens",9),
rep("Southwest Queens",8), rep("West Queens",7))

qns_data <- data.frame("ZipCode" = queens_zips,
"Neighborhood"=queens_neighborhoods)

statenisland_zips <- c(10302, 10303, 10310,	10306, 10307, 10308, 10309, 10312,
10301, 10304, 10305,10314)
statenisland_zips <- as.factor(statenisland_zips)

statenisland_neighborhoods <- c(rep("Port Richmond",3), rep("South Shore",5),
rep("Stapleton and St. George",3), "Mid-Island")

si_data <- data.frame("ZipCode" = statenisland_zips,
"Neighborhood"=statenisland_neighborhoods )

all_nyc_data <- rbind(si_data, qns_data, manh_data, bklyn_data, bronx_data)

```

Needed to download the ZipCode Tabulation file from here:
https://www.census.gov/geo/maps-data/data/cbf/cbf_zcta.html

```{r Mapping}

#install.packages("tmap")
#install.packages("leaflet")
library("tmap")
library("leaflet")

usshapefile <- "cb_2015_us_zcta510_500k.shp"
usgeo <- read_shape(file=usshapefile)

zip_codes_of_contribs_but_not_in_geo <- setdiff(fec_P2016$zip,
usgeo@data$ZCTA5CE10)
length(which(fec_P2016$zip %in% zip_codes_of_contribs_but_not_in_geo))

```

Interesting. There are 116 zip codes of contributors that dont have any 
geographic information. This amounts to 1905 contributions that can't be mapped 
without more information. [e.g. 10158, 10104, etc.]

Turns out after further research that some zip codes that are used by 
contributors are not 'official' zip codes. some of them are subsumed by other 
USPS codes.
See
http://newyork.hometownlocator.com/zip-codes/data,zipcode,10104.cfm
as an e.g; 10104 is contained within 10019

But I won't worry about these. In fact, if I do:

```{r}

setdiff(all_nyc_data$ZipCode, usgeo@data$ZCTA5CE10)
```

This shows that all but one zip code in the NYC zipcode data [in the
Far Rockaways] is in the geographic data, so I should be pretty good here.

And interestingly, 

```{r}

setdiff(fec_P2016$Zip, all_nyc_data$ZipCode)
setdiff(all_nyc_data$ZipCode, fec_P2016$zip)
```

you can see that there are only 2 zip codes in NYC that did not make any 
contributions at all [both in Queens]

```{r}

nystate_geo <- usgeo[usgeo@data$ZCTA5CE10 %in% all_nyc_data$ZipCode,]
qtm(nystate_geo)

```
<br>
Yup! That looks like all five boroughs!

```{r}

nyc_contributions <- subset(fec_P2016, zip %in% all_nyc_data$ZipCode)
nyc_contribs_by_full_zip <- nyc_contributions %>% group_by(zip) %>%
  summarise(sum_contrib = sum(contb_receipt_amt),n=n()) 

nymap <- append_data(nystate_geo, nyc_contribs_by_full_zip, 
key.shp ="ZCTA5CE10",key.data="zip")
  
qtm(nymap, "sum_contrib")

```
  
Nice static visualization of where the money is coming from. That's it for now!

#####

#Final Plots and Summary

###First Plot
```{r, fig.width=10, fig.height=7, echo=FALSE}

ggplot(data=contribs_by_zip, aes(x=zip_three_dig, y=sum_contrib)) + 
    geom_bar(stat="identity") + 
    ggtitle("Contribution Amounts by Zip Code [SCF]") +
    xlab("SCF Zip") + ylab("Contribution Amount ($)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_y_continuous(labels = comma)
```
    
###Description One

This plot compares the sum of contributions made per zip code to presidential
candidates (independent of candidates). Basically, it shows how much money the
population in each zip code across NY State gave to the candidates. I chose
this plot because it is a revelatory depiction of the disparity in contribution
amounts across NY state; there is an overwhelming concentration of money coming
from the 2 zip codes: 100 & 101. None of the other zip codes in New York State
come even close to matching the contributions made from just these two zip
codes. Furthermore, these two zip codes are based in Manhattan and a subset of
Manhattan at that. While it wasn't a surprise that Manhattan had the largest
dollar amount in contributions, it was surprising to see how concentrated it
was. There are a couple of weird zip codes such as [-11, 000, 003, 011] but
since the amounts barely registered, I did not pursue any further analysis on
them.

###Second Plot

```{r, fig.width=10, fig.height=7, echo=FALSE}

ggplot(data=new_d, aes(dat, cand_nm)) + geom_line() + scale_x_date(limits=as.Date(c(NA,'2016-04-01')),date_breaks="1 month") +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
geom_vline(aes(xintercept = as.numeric(as.Date("2016-01-01"))), linetype=2) +
geom_point(aes(drop_out_date, cand_nm, color='red')) +
ggtitle("Contribution dates by candidate") +
    xlab("Date") + ylab("Candidate") + guides(colour=FALSE)
```

###Description Two

This plot depicts the earliest and latest dates a candidate has received
contributions (the latest date being 31-March-2016 as that is where this
particular dataset ends). One can see that Marco Rubio started collecting
contributions as far back as 2013. The plot also shows all the candidates still
receiving contributions even on the last possible day. While that makes sense
for candidates who are still in the running (Hillary Clinton, Bernard Sanders,
Donald Trump), this seems odd for candidates who have dropped out. Why I picked
this chart as part of my final three is because of this oddity, i.e., it
reveals that candidates are still collecting money even after they dropped out!
For a short period of time after they drop out, collecting seems plausible 
because perhaps there is a pipeline of cash to be deposited, but Scott
Walker and Lindsey Graham dropped out in 2015 yet are still collecting money!
It definitely seems like something that requires more investigation.

###Third Plot

```{r, fig.width=10, fig.height=7, echo=FALSE}

ggplot(data=amounts_by_day, aes(contb_receipt_dt, amt, color=cand_nm)) +
    geom_line() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_y_continuous(labels = comma) + scale_x_date(limits=as.Date(c(NA,'2016-04-01')),date_breaks="1 month") + 
    ggtitle("Contributions to candidates by Date") +
    ylab("Amount Contributed ($)") + xlab("Date")
    

```


###Description Three

This graph shows the sum of contributions on a daily basis for all the
candidates from the inception of their campaigns until the last date in the
dataset (31-March-2016). Why I found this plot interesting is the huge spikes
that Hillary Clinton exhibits at the ends of the months. The spikes tower over
everyone else and indicate how strong her fundraising apparatus is (either that
or people are just enamored with her). The size of the spikes is one thing,
however, they also come at the end of the months. While this is definitely an
interesting phenomenon, I think it's readily explained by the huge end of monthly
drives that campaigns make to meet FEC monthly reporting periods. Anyone who
subscribes to political emails will have been subject to these.

#Reflection

The presidential campaign data set for New York from the FEC contained more
than 183,000 contributions ranging from 2013 until the end of March 2016. It
was interesting to see that candidates could solicit money from very early on 
(even after Obama was re-elected in 2012). However, I do have a faint 
recollection that a candidate couldn't go into real fundraising mode until 
after a declaration of candidacy. I recall Jeb Bush somehow building up a 
significant war chest as he was 'exploring a bid' but he hadn't yet declared 
his candidacy. More research into this would be required to make sense of this 
data and perhaps to draw sharper distinctions.


Further areas of analysis:

1. Break down which zips within the 100 & 101 SCF regions had the most 
contributions to further examine the concentration.

2. The occupations of various contributors are free-text and could be anything.
A lot of work can be done here to consolidate categories. For example, there 
was "Attorney" and "Attorney " [with a blank space] as well as "Lawyer". 
These records could all be merged. Unfortunately the largest set of
contributions came from an occupation of "". Who knows what these people do
for a living?

#Resources
http://stackoverflow.com/questions/13239639/duplicate-row-names-error-reading-table-row-names-null-shifts-columns/22408965#22408965

###how to tilt the x-axis labels so the candidate names can be read more clearly
http://stackoverflow.com/questions/15951216/too-many-factors-on-x-axis

### to set breaks on the x-axis for dates
http://stackoverflow.com/questions/14162829/set-date-range-in-ggplot

### help using aggregate function
http://davetang.org/muse/2013/05/22/using-aggregate-and-apply-in-r/

### help with not in operator
https://stat.ethz.ch/pipermail/r-help/2011-May/277501.html

### help with refactoring after subsetting data
http://stackoverflow.com/questions/27296310/refactor-whole-data-frame

### Help with Dplyr

http://stackoverflow.com/questions/35888508/r-dplyr-conditional-sum-with-dynamic-conditions

http://stackoverflow.com/questions/25571547/select-unique-values-with-select-function-in-dplyr-library

http://stackoverflow.com/questions/3777174/plotting-two-variables-as-lines-using-ggplot2-on-the-same-graph

### Zip Code download
http://proximityone.com/cen2010_zcta_dp.htm

### Information on significance of 3 digit zip code so treating this as 1 group
### Sectional Center Facility [SCF]

http://www.zipboundary.com/zipcode_faqs.html

###Info on ColClasses option in read.csv
http://stats.stackexchange.com/questions/1471/is-it-possible-to-directly-read-csv-columns-as-categorical-data